---
title: "dashboard_sleep"
output: html_document
date: "`r base::date()`"
chunk_output_type: console
---

```{r knit into synapse, eval=FALSE, include=FALSE}
## New lines for calling synapseclient
synapseclient <- reticulate::import("synapseclient")
syn_temp <- synapseclient$Synapse()
syn_temp$login() # is fails use syn_temp$login(authToken = <yourPAT>)
# get your PAT at "https://www.synapse.org/#!PersonalAccessTokens:"

### Note the new function being called from knit2synapse
### Please install this version of knit2synapse
### devtools::install_github("jgockley62/knit2synapse")

### Please install v1.28 of reticulate, as this is the needed version for synapser
### devtools::install_version("reticulate", version = "1.28", repos = "http://cran.us.r-project.org")

library(synapser)
library(knit2synapse)
synapser::synLogin()

knit2synapse::createAndKnitToFolderEntityClient(file = "dashboard_sleep_archive_markdown.Rmd",
                                          parentId = "syn52540337",
                                          folderName = "dashboard_sleep_archive_test")
```

```{r required libraries, include=FALSE, warning=FALSE, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE)
########
# Required Libraries
########
library(synapser)
library(tidyverse)
library(arrow)
# library(lubridate)
library(knitr)
```

## Sleep Metrics Dashboard (fitbit_sleeplogs)

Summary metrics and visualizations of data derived from the dataset - fitbit_sleeplogs

```{r setup and download, include=FALSE, warning=FALSE, message=FALSE}

synapser::synLogin()
ARCHIVE_VERSION <- '2023-09-21'

## Set up Token access
sts_token <- synapser::synGetStsStorageToken(entity = 'syn52506069', # sts enabled destination folder
                                             permission = 'read_only',   # request a read only token
                                             output_format = 'json')

s3_external <- arrow::S3FileSystem$create(access_key = sts_token$accessKeyId,
                                          secret_key = sts_token$secretAccessKey,
                                          session_token = sts_token$sessionToken,
                                          region="us-east-1")

## Get list of datasets in the S3 bucket
base_s3_uri_external <- paste0(sts_token$bucket, "/", sts_token$baseKey,'/',ARCHIVE_VERSION)
parquet_datasets_external <- s3_external$GetFileInfo(arrow::FileSelector$create(base_s3_uri_external, recursive=F))

## Get all valid datasets
i <- 0
valid_paths <- character()
for (dataset in parquet_datasets_external) {
  if (grepl('recover-main-project/main/archive/', dataset$path, perl = T, ignore.case = T)) {
    i <- i+1
    cat(i)
    cat(":", dataset$path, "\n")
    valid_paths <- c(valid_paths, dataset$path)
  }
}

## Get dataset type (for eg., dataset_enrolledparticipants)
valid_paths_ext_df <- valid_paths %>% 
  as.data.frame() %>% 
  `colnames<-`('parquet_path_external') %>% 
  dplyr::rowwise() %>% 
  dplyr::mutate(datasetType = str_split(parquet_path_external,'/')[[1]][5]) %>%
  dplyr::ungroup()

#### Get the sleep data in
dataset_path <- valid_paths[[9]]
fitbit_sleeplogs <-  arrow::open_dataset(s3_external$path(as.character(dataset_path))) %>% dplyr::collect()

dataset_path <- valid_paths[[1]]
enrolledparticipants <-  arrow::open_dataset(s3_external$path(as.character(dataset_path))) %>% dplyr::collect()
```


```{r needed metrics, include=FALSE, message=FALSE, warning=FALSE, echo=FALSE}
##############
### fitbit_sleeplogs
## Summarizations needed:
## (Sleep - Classic or stages)
#   1. Number of Sleep Records
#   2. Number of Sleep participants
#   3. Number of Sleep days
#   4. Average Number of Sleep days per participant
#   5. Proportion of Sleep days since enrollment
## (Sleep - Stages)
#   6. Number of Sleep Stages participants
#   7. Number of Sleep Stages days
#   8. Average Number of Sleep Stages days per participant
#   9. Proportion of Sleep Stages days since enrollment
##############
```

### Number of records

```{r Number of records, echo=FALSE, warning=FALSE, message=FALSE}
# 1 
number_of_records <- fitbit_sleeplogs %>% 
  dplyr::group_by(Type) %>% 
  dplyr::count() %>% 
  dplyr::full_join(data.frame(Type = 'Total',
                              n = nrow(fitbit_sleeplogs)))

kable(number_of_records)
```

### Number of participants
```{r Number of participants, echo=FALSE, warning=FALSE, message=FALSE}
# 2 and 6
number_of_sleep_participants <- fitbit_sleeplogs %>% 
  dplyr::select(ParticipantIdentifier, Type) %>% 
  unique() %>% 
  dplyr::left_join(fitbit_sleeplogs %>% 
                     dplyr::select(ParticipantIdentifier, Type) %>% 
                     unique() %>% 
                     dplyr::group_by(ParticipantIdentifier) %>%
                     dplyr::count() %>% 
                     dplyr::ungroup()
  ) %>% 
  dplyr::rowwise() %>% 
  dplyr::mutate(newType = ifelse(n==2,'classic and stages',Type)) %>% 
  dplyr::ungroup() %>% 
  dplyr::select(-Type) %>% 
  unique() %>% 
  dplyr::group_by(newType) %>% 
  dplyr::count() %>% 
  dplyr::ungroup()

number_of_sleep_participants <- rbind(number_of_sleep_participants, data.frame('newType' = 'Total', 'n' = sum(number_of_sleep_participants$n)))

number_of_sleep_participants <- number_of_sleep_participants %>% 
  dplyr::rename(type = newType)

kable(number_of_sleep_participants)
```

### Number of Sleep days

Assuming that a record that is atleast 3 hrs long to be considered a sleep record. A day with atleast one sleep record is considered a sleep day
```{r Number of Sleep days, echo=FALSE, warning=FALSE, message=FALSE}
# 3 and 7 (consider atleast 3 hrs for a sleep day)
number_of_sleep_days <- fitbit_sleeplogs %>% 
  dplyr::select(ParticipantIdentifier, StartDate, EndDate, Duration, Type) %>% 
  dplyr::rowwise() %>% 
  dplyr::mutate(dayOfRecording= as.Date(StartDate),
                Duration = as.numeric(Duration)) %>% 
  dplyr::filter(Duration >= 180) %>% # atleast 180min ie 3Hrs to be considered a sleep day/record
  dplyr::ungroup() %>% 
  dplyr::select(ParticipantIdentifier, dayOfRecording, Type) %>%
  unique() 

number_of_sleep_days <- number_of_sleep_days %>% 
  dplyr::left_join(number_of_sleep_days %>% 
                     dplyr::group_by(ParticipantIdentifier, dayOfRecording) %>% 
                     count()) %>% 
  dplyr::mutate(newType = ifelse(n==2, 'classic and stages', Type)) %>% 
  dplyr::select(-Type) %>% 
  unique() %>% 
  dplyr::group_by(ParticipantIdentifier, newType) %>% 
  count() %>% 
  dplyr::rename(Type = newType) %>% 
  dplyr::ungroup()

options(repr.plot.width = 9, repr.plot.height =6)
ggplot2::ggplot(data = number_of_sleep_days, aes(x = n))+ geom_histogram() + xlab('Number of days with atleast one 3 hr record (log scale)') + ylab('Number of records') +
  theme_minimal() + facet_wrap(~Type) + scale_x_log10()

```

### Average number of sleep days
A day might have both classic and stages Type sleep records for a given participant. We are counting both in the averages below

```{r Avg number of sleep days, echo=FALSE, warning=FALSE, message=FALSE}
# 4 and 8
avg_number_of_sleep_days <- fitbit_sleeplogs %>% 
  dplyr::select(ParticipantIdentifier, StartDate, EndDate, Duration, Type) %>% 
  dplyr::rowwise() %>% 
  dplyr::mutate(dayOfRecording= as.Date(StartDate),
                Duration = as.numeric(Duration)) %>% 
  dplyr::filter(Duration >= 180) %>% # atleast 180min ie 3Hrs to be considered a sleep day/record
  dplyr::ungroup() %>% 
  dplyr::select(ParticipantIdentifier, dayOfRecording, Type) %>%
  unique() %>% 
  dplyr::group_by(ParticipantIdentifier, Type) %>% 
  dplyr::count() %>% 
  dplyr::ungroup() %>% 
  dplyr::group_by(Type) %>% 
  dplyr::summarise(mean = mean(n),
                   median = median(n))

kable(avg_number_of_sleep_days)
```


### Proportion of sleep days since enrollment
```{r Proportion of sleep days since enrollment,  echo=FALSE, warning=FALSE, message=FALSE}
# 5 (Propotion of sleep days since enrollment)
sleep_days_since_enrollment <- fitbit_sleeplogs %>% 
  dplyr::left_join(enrolledparticipants %>% 
                     dplyr::select(ParticipantIdentifier, EnrollmentDate) %>% 
                     dplyr::mutate(days_enrolled = lubridate::days(Sys.Date() - as.Date(EnrollmentDate)))) %>% 
  dplyr::rowwise() %>% 
  dplyr::filter(StartDate >= EnrollmentDate) %>%  # records after enrollment
  dplyr::rowwise() %>% 
  dplyr::mutate(dayOfRecording= as.Date(StartDate),
                Duration = as.numeric(Duration)) %>% 
  dplyr::filter(Duration >= 180) %>% # atleast 180min ie 3Hrs to be considered a sleep day/record
  dplyr::ungroup() %>% 
  dplyr::select(ParticipantIdentifier, dayOfRecording, days_enrolled, Type) %>%
  unique() %>% 
  dplyr::group_by(ParticipantIdentifier, days_enrolled, Type) %>% 
  count() %>%
  dplyr::ungroup() %>% 
  dplyr::rename(ndays_after_enrollment = n) %>% 
  dplyr::rowwise() %>% 
  dplyr::mutate(proportion_sleep_days_post_enroll = 100*ndays_after_enrollment/lubridate::day(days_enrolled)) %>% 
  dplyr::ungroup()

options(repr.plot.width = 8, repr.plot.height =6)
ggplot2::ggplot(data = sleep_days_since_enrollment, aes(x = proportion_sleep_days_post_enroll))+ geom_histogram() + xlab('Proportion of Sleep days Since enrollment') + ylab('Number of records') + facet_wrap(~Type)+
  theme_minimal()

mean_proportion_sleep_days_since_enrollment <- sleep_days_since_enrollment %>% 
  dplyr::group_by(Type) %>% 
    dplyr::summarise(mean = mean(proportion_sleep_days_post_enroll),
                   median = median(proportion_sleep_days_post_enroll))
kable(mean_proportion_sleep_days_since_enrollment)
```


### Average number of sleep days after enrollment
A day might have both classic and stages Type sleep records for a given participant. We are counting both in the averages below

```{r Avg number of sleep days after enrollment, echo=FALSE, warning=FALSE, message=FALSE}
# 4 and 8
avg_number_of_sleep_days_enrollment <- sleep_days_since_enrollment %>% 
  dplyr::group_by(Type) %>% 
  dplyr::summarise(mean = mean(ndays_after_enrollment),
                   median = median(ndays_after_enrollment))

kable(avg_number_of_sleep_days_enrollment)
```
